---
title: "Untitled"
author: "MacKenzie Ullman"
date: "2/9/2021"
output: pdf_document
---

Import the “beach.csv” dataset into R and examine the first few rows of data
```{r}
setwd("F:/Quant Eco")
getwd()
beach <- read.csv(file='beach.csv')
head(beach)

```

Fit a model that includes interactions between Year and BeachID. Report estimates of all coefficients
```{r}
fit_i <- lm(OpenBeach ~ BeachID * Year, data = beach)
fit_i

```

Predict the beach area at all three beaches in 1992 
```{r}
betas <- coef(fit_i)
betas
pred1 <- data.frame(BeachID=c('A'), Year=c(1992))
predict(fit_i, newdata = pred1)
```
The predicted beach area for beach A in 1992 is 362.815 hectares.

```{r}
pred2 <- data.frame(BeachID=c('B'), Year=c(1992))
predict(fit_i, newdata = pred2)
```
The predicted beach area for beach B in 1992 is 100.034 hectares.

```{r}
pred3 <- data.frame(BeachID=c('C'), Year=c(1992))
predict(fit_i, newdata = pred3)
```
The predicted beach area for beach C in 1992 is 214.3744 hectares.


Plot the predicted area of each beach across all years 
```{r}
x <- 1992:2014
b_0 <- 36945.72537
b_b1 <- -34211.97897
b_b2 <- -28776.11491
b_b3<- -18.36491
b_b4 <- 17.04277
b_b5<- 14.37132
y_ba <- b_0 + b_b3*x 
y_ba
y_bb <- b_0 + b_b1*1 + b_b3*x + b_b4*x
y_bb
y_bc <-b_0 + b_b2*1 + b_b3*x + b_b5*x 
y_bc
```
```{r}
betas
```
```{r}
plot(x=x, y=y_ba, type= 'b', xlab= 'Year', ylab= 'Area', col='black')
lines(x=x, y=y_bb, col='red')
lines(x=x, y=y_bc, col='blue')
```


Interpret the rate of beach area change at each of the 3 beaches 
```{r}
ba<-b_b3
ba
bb<-b_b3+b_b4
bb
bc<-b_b3+b_b5
bc

```

The rate of beach area change at beach A is a decrease by 18.36491 hectares per year.
The rate of beach area change at beach B is a decrease of 1.32214 hectaeres per year.
The rate of beach area change at beach C is a decrease of 3.99359 hectares per year. 