---
title: "Homework 7"
author: "MacKenzie Ullman"
date: "4/13/2021"
output: word_document
---

1. Load data and place into an unmarkedFrameOccu object
```{r}
setwd("F:/Quant Eco")
getwd()
HW7 <- read.csv(file ='Bobcat.csv')
head(HW7)
summary(HW7)
```
```{r}
library(unmarked)
bobcat_mat <- as.matrix(HW7)
nmix_data <- unmarkedFramePCount(y = bobcat_mat)

```

```{r}
#detection covariates
p_covs <- read.csv('p covariates.csv')
head(p_covs)

class(p_covs)

det_covs <- list(
people = data.frame(p_covs[, c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71)])
)

nmix_data <- unmarkedFramePCount(y = as.matrix(bobcat_mat), obsCovs = det_covs)


#site level covariates
site_covs <- read.csv('psi covariates.csv')
head(site_covs)

class(site_covs)

nmix_data <- unmarkedFramePCount(y = as.matrix(bobcat_mat), siteCovs = site_covs, obsCovs = det_covs)

```


2. Fit the following candidate set of models:
```{r}
library(unmarked)
#intercept, intercept model
fit1_p <- pcount(formula = ~ 1 ~ 1 , data = nmix_data, K = 100, mixture = "P")
summary(fit1_p)

#people, intercept
fit2_p <- pcount(formula = ~ people ~ 1 , data = nmix_data, K = 100, mixture = "P")
summary(fit2_p)

#intercept, disturbance
fit3_p <- pcount(formula = ~ 1 ~ Dist_5km , data = nmix_data, K = 100, mixture = "P")
summary(fit3_p)

# people disturbance
fit4_p <- pcount(formula = ~ people ~ Dist_5km , data = nmix_data, K = 100, mixture = "P")
summary(fit4_p)

```

```{r}
#intercept, intercept model
fit1_nb <- pcount(formula = ~ 1 ~ 1 , data = nmix_data, K = 100, mixture = "NB")
summary(fit1_nb)

#people, intercept
fit2_nb <- pcount(formula = ~ people ~ 1 , data = nmix_data, K = 100, mixture = "NB")
summary(fit2_nb)

#intercept, disturbance
fit3_nb <- pcount(formula = ~ 1 ~ Dist_5km , data = nmix_data, K = 100, mixture = "NB")
summary(fit3_nb)

# people disturbance
fit4_nb <- pcount(formula = ~ people ~ Dist_5km , data = nmix_data, K = 100, mixture = "NB")
summary(fit4_nb)
```

3.Perform model selection with AIC. What is your top model? How do you know? Is there model selection uncertainty?
```{r}
library(AICcmodavg)
```
```{r}
cand.set <- list(
P1 = fit1_p, P2 = fit2_p, P3 = fit3_p, P4 = fit4_p,
NB1 = fit1_nb, NB2 = fit2_nb, NB3 = fit3_nb, NB4 = fit4_nb
)

mods <- aictab(cand.set = cand.set, second.ord = F)
head(mods)
```
The top model is NB3 (fit3_nb <- pcount(formula = ~ 1 ~ Dist_5km , data = nmix_data, K = 100, mixture = "NB")). This is the top model because it is at the top of the tabla and it  has the lowest AIC score of 4405.99. There is a little uncertainty because the delta AIC of the second top model is less than 2. However there isnt much uncertainty because there is only  one other model with a delta AIC score less than 2 when compared to the top model Delta AIC, not multiple models with a delta AIC score less than 2. 


4. Average both the effect of people on detection, and disturbance on occupancy, over all models. Report model-averaged slope coefficients and 95% confidence intervals.
```{r}
#effect of people on detection
avg_people_1 <- modavgShrink(cand.set = cand.set,
parm = 'people',
second.ord = F,
parm.type = 'detect')

avg_people_1$Mod.avg.beta

avg_people_1
```
The model averaged coefficient of people on detection is -0.02583
The 95% confidence interval is -0.42, 0.36.

```{r}
#effect of disturbance on occupancy
avg_disturbance_1 <- modavgShrink(cand.set = cand.set,
parm = 'Dist_5km',
second.ord = F,
parm.type = 'lambda')

avg_disturbance_1$Mod.avg.beta

avg_disturbance_1
```
The model averaged slope coefficient of disturbance on occupancy is -27.66347
The 95% confidence interval is -37.27, -18.06. 


5. Obtain and plot model-averaged predictions of occupancy probability and detection probability. Average over all models, and make predictions over the observed range of each variable.
```{r}
# data.frame for new predictions
new_dat <- data.frame(
Dist_5km = seq(from = min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100),
people = seq(from = 1, to = 71, length.out = 100)
)

# calculating model averaged predictions
#detection probability
avg_prd <- modavgPred(cand.set = cand.set,
newdata = new_dat,
second.ord = F,
parm.type = 'detect')
avg_prd
#occupancy probability
avg_prd2 <- modavgPred(cand.set = cand.set,
newdata = new_dat,
second.ord = F,
parm.type = 'lambda')
avg_prd2


```


