---
title: "Homework 5"
author: "MacKenzie Ullman"
date: "4/5/2021"
output: pdf_document
---
1. Load data and place into an unmarkedFrameOccu object
```{r}
setwd("F:/Quant Eco/")
getwd()
y <- read.csv(file= 'Bobcat.csv')
head(y)
summary(y)
```
```{r}
install.packages('unmarked')
library(unmarked)
```

```{r}
bobcat_mat <- as.matrix(y)
occu_data <- unmarkedFrameOccu(y = bobcat_mat)
occu_data
```



2. Fit an occupancy model that assumes conditional detection probability is a function of the number of people photographed at a site, and occupancy probability is a function of disturbance within 5km.


```{r}
p_covs <- read.csv('p covariates.csv')
head(p_covs)

class(p_covs)

det_covs <- list(
people = data.frame(p_covs[, c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71)])
)

occu_data <- unmarkedFrameOccu(y = bobcat_mat, obsCovs = det_covs)





site_covs <- read.csv('psi covariates.csv')
head(site_covs)

class(site_covs)

occu_data <- unmarkedFrameOccu(y = bobcat_mat, siteCovs = site_covs, obsCovs = det_covs)

fit <- occu(formula = ~ people ~ Dist_5km , data = occu_data)
summary(fit)
```


3. Interpret the effect of the number of people photographed at each site on conditional detection probability of bobcat. Verity your interpretation in R.

For every one change in the number of people photographed, the log odds of detecting a bobcat at a camera decreases by 0.0608. 
```{r}
beta <- coef(fit)
beta

a <- beta[4] * 1
a

b <- beta[4] * 2
b

c <- b-a
c
```

4. Re-create the Wald test for the effect of disturbance within 5km. What is the null hypothesis? Do you reject or fail to reject this null hypothesis?
```{r}
w <- -23.643839/4.7728768
w

2 * pnorm(-1 * abs(w))
```
We reject the null hypothesis


5. Predict and plot the effect of disturbance within 5km on bobcat occupancy probability. Do this over the range of observed disturbance within 5km.

```{r}
new <- data.frame(disturbance = seq(from = min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100))

predict <- predict(object= fit, newdata = new, type = 'state')
```

```{r}
x1 <- seq(from = min(site_covs$Dist_5km), to = max(site_covs$Dist_5km), length.out = 100)
plot(y= predict(y), x= x1, ylab= 'Bobcat Occupancy Probability' , xlab= 'Disturbance Within 5km', type= 'l' )
```

