---
title: "Homework 3"
author: "MacKenzie Ullman"
date: "3/2/2021"
output: pdf_document
---

Fit a logistic regression model that assumes the probability of success is an additive function of variables
x1 and x2.

```{r}
setwd("F:/Quant Eco")
getwd()
HW3 <- read.csv(file ='HW3.csv')
head(HW3)
summary(HW3)
```

```{r}
fit <- glm( y~ x1+x2, family= binomial, data= HW3)
summary(fit)
```

Interpret the effect of variable x1 on the log odds of success. Verify your interpretation in R.

when x1 changes by 1 unit, the log odds of success change by -0.1679
```{r}
betas <- coef(fit)
betas

yx<- betas[1]+betas[2]*0
yx

yx1 <-betas[1]+betas[2]*1
yx1

yx1-yx
```


Interpret the effect of variable x2 on the log odds of success. Verify your interpretation in R.

when the x2 variable changes by one unit, the log odds of success change by -0.9679
```{r}
betas <- coef(fit)
betas

yx2 <- betas[1] + betas[3]*0
yx2

yx21 <- betas [1] + betas[3]*1
yx21

yx21-yx2
```


Duplicate the Wald Test and p-values for variables x1 and x2 performed by the glm() function. Do you reject or fail to reject your null hypothesis?
```{r}
#wald test for x1
betas[2] / summary(fit)[['coefficients']]['x1', 'Std. Error']
summary(fit)[['coefficients']]['x1', 'z value']
```

```{r}
#wald test for x2
betas[3] / summary(fit)[['coefficients']]['x2b', 'Std. Error']
summary(fit)[['coefficients']]['x2b', 'z value']
```
```{r}
# p-values
#test statistic for x1
tsx1 <- betas[2] / summary(fit)[['coefficients']]['x2b', 'Std. Error']
tsx1
#test statistic for x2
tsx2 <- betas[3] / summary(fit)[['coefficients']]['x2b', 'Std. Error']
tsx2
#p-value for x1
2 * pnorm(-1 * abs(tsx1), mean = 0, sd = 1)
#p-value for x2
2 * pnorm(-1 * abs(tsx2), mean = 0, sd = 1)
```

We reject the null hypothesis for x1. We fail to reject the null hypothesis for x2. 

Predict and plot the mean probability of success over the range of values of x1.
```{r}
plogis(yx2)

x1 <- seq(from = min(HW3$x1), to = max(HW3$x1),length.out = 100)

y <- betas[1] + betas[2] * x1 
plot(x = x1, y = plogis(y), ylab = 'Probability of Success over Range of x1',
xlab = 'x1', cex.axis = 1.5, cex.lab = 1.5, type = 'l')
```


