---
title: "Homework 6"
author: "MacKenzie Ullman"
date: "4/9/2021"
output: pdf_document
---
1. Load data and place into an unmarkedFramePCount object
```{r}
setwd("F:/Quant Eco")
getwd()
HW6 <- read.csv(file ='count.csv')
head(HW6)
summary(HW6)
```
```{r}
library(unmarked)

count_mat <- as.matrix(HW6)
nmix_data <- unmarkedFramePCount(y = count_mat)
```


2. Fit an N-mixture model that assumes conditional detection probability is a function of the detection covariate provided, and expected abundance is an additive function of variables x1 and x2.
```{r}
#detection covariates
p_covs <- read.csv('obs_covs.csv')
head(p_covs)

#Placing detection covariates in an unmarkedFramePCount object
det_covs <- list(
replicate = data.frame(p_covs[, c('j1', 'j2', 'j3')])
)

#Placing the list of detection covariates in to the unmarkedFramePCount object
nmix_data <- unmarkedFramePCount(y = as.matrix(count_mat), obsCovs = det_covs) 

fit <- pcount(formula = ~ replicate ~ 1, data = nmix_data, K = 100)
```

```{r}
#site level covariates
sitecovs <- read.csv('site_covs.csv')
head(sitecovs)


nmix_data <- unmarkedFramePCount(y = as.matrix(count_mat), siteCovs = sitecovs, obsCovs = det_covs) 

fit <- pcount(~ replicate ~ x1 + x2,data = nmix_data, K = 100)
summary(fit)
```

3. Interpret the effect of x1 on the expected count at each site. Verity your interpretation in R.
The expected count per each site increases by 0.3696188 when x1 increases by 1 unit.
```{r}
#verify
beta <- coef(fit)
beta

a <- beta[2]*1
a

b <- beta[2]*2
b

c <- b-a
c
```

4. Predict and plot the effect of the supplied detection covariate. Do this over the range of this covariate.
```{r}
new <- data.frame(replicate = seq(from = min(det_covs$replicate), to = max(det_covs$replicate), length.out = 100))


prd <- predict(object = fit, newdata = new, type = 'det')

prd

```

```{r}
plot(x = c(0,0), y = prd[1, c('lower', 'upper')],

ylab = 'Detection probability', xlab = '',)

```
5. Use contrasts to compare expected abundance between all pairwise levels of variable x2. Obtain p-values associated with each contrast and tell me whether you reject or fail to reject each null hypothesis tested.
```{r}
x <- matrix(
c(0, 0, 1, -1, 0,
0, 0, 1, 0, -1,
0, 0, 0, 1, -1),
nrow = 3, byrow = T
)
x

lin_com <- linearComb(obj = fit, coefficients = x, type = 'state')

lin_com
```
```{r}
w <- coef(lin_com) / SE(lin_com)
w

#Calculating p-values
2 * pnorm(-1 * abs(w))
```
We reject all the null hypotheses. There is a difference between b and c, a difference between b and d, and a difference between c and d,  in terms of abundance probability
