---
title: "Homework 4"
author: "MacKenzie Ullman"
date: "3/15/2021"
output: pdf_document
---

```{r}
setwd("F:/Quant Eco/")
getwd()
hw4 <- read.csv(file= 'Homework 4 Data.csv')
head(hw4)
summary(hw4)
```

Fit a Poisson regression model that assumes expected count is an interactive function of variables x1 and x2.
```{r}
fit <- glm( y ~ x1 * x2, family= poisson, data= hw4)
summary(fit)
```

Interpret the effect of variable x1 on the expected count when x2 is fixed at level "b". Verify your interpretation in R.

```{r}
beta <- coef(fit)
beta

x2 <- 1
c_1<- beta[2] + beta[4] * 1
c_1
```

```{r}
exp(c_1)

# Lambda increases by 46% for every 1 unit increase in x1
```
```{r}
#verify
c_2 <- beta[2] + beta[4] * 2
c_2

exp(c_2)

exp(c_2) / exp(c_1)
exp(beta[4])

```

Interpret the effect of variable x2 on the expected count when x1 is fixed at 1. Verify your interpretation
in R.
```{r}
beta
x1 <- 1
c_x2 <- beta[3] + beta[4] * 1
c_x2
```
```{r}
exp(c_x2)

# Lambda increases by 56% for every 1 unit increase in x2
```
```{r}
#verify
c2_x2 <- beta[3] + beta[4] * 2
c2_x2

exp(c2_x2)

exp(c2_x2)/ exp(c_x2)
exp(beta[4])
```


Predict the expected count, ± 95% confidence intervals, over the observed range of values of x1, assuming x2 is fixed at level "b".

```{r}
# x1 range of values
x1_new <- data.frame(
  x1 = seq(min(hw4$x1), max(hw4$x1), length.out = 100),
  x2 = factor(rep('b', times= 100), levels = c('a','b')))

# predicted expected count
prd <- predict.glm(object= fit, newdata = x1_new, type = 'link', se.fit = T)
prd

#confidence intervals
low <- exp(prd$fit - qnorm(.975) * prd$se.fit)
high <- exp (prd$fit + qnorm(.975) * prd$se.fit)

# plot
plot(y = exp(prd$fit), x = x1_new$x1, xlab = 'x1',
ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
ylim = c(min(low), max(high)), type = 'l')
lines(x = x1_new$x1, y = low, lty = 2)
lines(x = x1_new$x1, y = high, lty = 2)
```

Predict the expected count, ± 95% confidence intervals, of levels "a" and "b", assuming x1 is fixed at it’s mean.
```{r}

x1_new2 <- data.frame(
  x1a = mean(hw4$x1),
  x2a = factor('a', levels = c('a','b')))
# x1 is mean and x2 is b
x1_new2b <- data.frame(
  x1b = mean(hw4$x1), 
  x2b = factor('b', levels = c('a','b')))


# predicted expected count
prd2a <- predict.glm(object= fit, newdata = x1_new2, type = 'link', se.fit = T)
prd2a

prd2b <- predict.glm(object= fit, newdata = x1_new2b, type = 'link', se.fit = T)
prd2b
#confidence intervals
low <- exp(prd2a$fit - qnorm(.975) * prd2a$se.fit)
high <- exp (prd2a$fit + qnorm(.975) * prd2a$se.fit)

low <- exp(prd2b$fit - qnorm(.975) * prd2b$se.fit)
high <- exp (prd2b$fit + qnorm(.975) * prd2b$se.fit)

# plot
plot(y = exp(prd2a$fit), x = x1_new2$x1, xlab = 'x1',
ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
ylim = c(min(low), max(high)), type = 'l')
lines(x = x1_new2$x1, y = low, lty = 2)
lines(x = x1_new2$x1, y = high, lty = 2)

plot(y = exp(prd2b$fit), x = x1_new2b$x1, xlab = 'x1',
ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
ylim = c(min(low), max(high)), type = 'l')
lines(x = x1_new2b$x1, y = low, lty = 2)
lines(x = x1_new2b$x1, y = high, lty = 2
```

